# Reproducible Research: Peer Assessment 1


## Introduction

It is now possible to collect a large amount of data about
personal movement using activity monitoring devices such as
a
[Fitbit](http://www.fitbit.com),
[Nike Fuelband](http://www.nike.com/us/en_us/c/nikeplus-fuelband),
or
[Jawbone Up](https://jawbone.com/up).
These type of devices are part of the "quantified self"
movement -- a group of enthusiasts who take measurements about
themselves regularly to improve their health, to find patterns
in their behavior, or because they are tech geeks. But these
data remain under-utilized both because the raw data are hard
to obtain and there is a lack of statistical methods and
software for processing and interpreting the data.

This assignment makes use of data from a personal activity
monitoring device. This device collects data at 5 minute
intervals through out the day. The data consists of two months
of data from an anonymous individual collected during the
months of October and November, 2012 and include the number of
steps taken in 5 minute intervals each day.


## Loading and preprocessing the data

The data for this assignment can be downloaded from the course web site:

* Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) [52K]

The variables included in this dataset are:

* **steps**: Number of steps taking in a 5-minute interval (missing
    values are coded as `NA`)

* **date**: The date on which the measurement was taken in YYYY-MM-DD
    format

* **interval**: Identifier for the 5-minute interval in which
    measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there
are a total of 17,568 observations in this dataset.

First, read the activity dataset:
```{r}
dataset_file <- unzip("activity.zip")
activity_data <- read.csv("activity.csv")
````

Next, transform the date factors to objects of the Date class:
```{r}
activity_data$date <- as.Date(activity_data$date)
```


## What is mean total number of steps taken per day?

For this part of the assignment, we will ignore the missing values (NA) in the activity dataset. To determine the total number of steps taken each day, we will use the aggregate function:
```{r}
steps_per_day <- aggregate(steps ~ date, data = activity_data, sum)
```

We can now plot a histogram showing the total number of steps taken each day:

```{r}
hist(steps_per_day$steps, breaks = 30, main = "Total number of steps taken each day", xlab = "Total number of steps")
```

The **mean** total number of steps taken per day is:
```{r}
mean(steps_per_day$steps)
```

The **median** total number of steps taken per day is:
```{r}
median(steps_per_day$steps)
```


## What is the average daily activity pattern?

1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

Again, we make use of the aggregate function:
```{r}
mean_steps <- aggregate(steps ~ interval, data = activity_data, mean)
plot(steps ~ interval, data = mean_steps, type = "l", xlab = "Interval", ylab = "Average number of steps taken")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_steps <- max(mean_steps$steps)
max_row <- which(mean_steps$steps == max_steps)
max_row
mean_steps[row(mean_steps) == max_row]
```

Interval **`r mean_steps[row(mean_steps) == max_row][1]`** contains the maximum number of steps **`r mean_steps[row(mean_steps) == max_row][2]`**.


## Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as `NA`). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. The total number of missing values in the dataset (i.e. the total number of rows with `NA`s) is:
```{r}
length(which(is.na(activity_data$steps)))
```

2. To fill in all of the missing values in the dataset, we will use...

## Are there differences in activity patterns between weekdays and weekends?

